Diversification of salmon fishermen within strategies (permits)

### 2016-08-11 salmon/analysis/salmon_paper.Rmd
Cleaned up figures

```{r, echo = FALSE, cache=TRUE, eval = FALSE}
# Effect of Diversification Across Space

#Using the same strategies as paper # 1, the comparisons across areas are mostly related to #salmon fisheries (these are areas with 2 or more strategies present and strategies with 2 or #more areas represented). And as in paper # 1, this is for people making >= 5K / year.

dat = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
### Eric's culling:
# Filters: remove people-year combinations making < $10000
dat = dat[-which(dat$nregion> 1 | dat$nregion.prev > 1 | (dat$region!=dat$region.prev)),]
dat$region[which(dat$region%in%c("SEAK Outside","SEAK Inside"))] = "SEAK"
dat = dat[dat$region%in%names(rev(sort(table(dat$region)))[1:10]), ]

dat = dat[which(dat$revenue >= 5000 & dat$revenue.prev >= 5000), ]

# note: grouping here is based on strategies defined by permits
dat$strategy = dat$strategy_permit
dat <- group_by(dat, strategy) %>%
  mutate(npeople = length(unique(p_holder)))

dat[dat$length<15 & !is.na(dat$length), ]$length <- NA

dat <- dat %>% group_by(boat) %>%
  mutate(median_length_boat = median(length, na.rm = T))

dat <- dat %>% group_by(strategy) %>%
  mutate(median_length_permit = median(length, na.rm = T))

dat <- dat %>% mutate(length = ifelse(is.na(length), median_length_boat, length))
dat <- dat %>% mutate(length = ifelse(is.na(length), median_length_permit, length))

#2. For each person-year combination, we created 'strategies' by concatenating
# all permits fished, and only retaining 'strategies' with >= 200 data points. Previously,
# this was done by people-year, but Eric changed this to be including > 200 people / strategy
top_strategies = group_by(dat, strategy) %>%
  summarize(n = length(unique(p_holder)),
    earn=sum(revenue)) %>%
  filter(n>=50) # note -- the code should work fine with this >= 100 too

# A few of these can be deleted (small in river fisheries, experimental, etc)
top_strategies = top_strategies[-which(top_strategies$strategy %in% c("S04X","S04Y","Z12B","R18B")),]

#3. We then tabulated the permits that make up these strategies, and there are only
# 56.
top_permits = data.frame("orig"=names(table(unlist(lapply(top_strategies$strategy, strsplit, " ")))))

# 4. of these top permits, we proceeded to group permits that were targeting a single species
# and other than area, were otherwise the same. In other words, someone fishing herring roe
# in kodiak (G34K) and alaska peninsula (G34M) have a herring roe strategy. This combining
# reduces permits to 59
top_permits$new = as.character(top_permits$orig)
top_permits$new[which(substr(top_permits$new,1,3)%in%c("D09","D9C","D9D"))] = "D09" # dungeness
top_permits$new[which(substr(top_permits$new,1,3)=="G01")] = "G01" # herring roe, purse seine
top_permits$new[which(substr(top_permits$new,1,3)=="G34")] = "G34" # herring roe, gillnet
top_permits$new[which(substr(top_permits$new,1,3)=="K91")] = "K91" # king crab
top_permits$new[which(substr(top_permits$new,1,3)=="L12")] = "L12" # herring, hand pick
top_permits$new[which(substr(top_permits$new,1,3)=="L21")] = "L21" # herring, pound
top_permits$new[which(substr(top_permits$new,1,3)=="P09")] = "P09" # shrimp pot gear
top_permits$new[which(substr(top_permits$new,1,3)=="Q11")] = "Q11" # sea cucumber
top_permits$new[which(substr(top_permits$new,1,3)=="T09")] = "T09" # tanner crab

# 5. Combining multispecies finfish permits (M). There are 8 miscellaneous finfish permits (M)
# remaining, and only 2 can be combined: M26B/G (mechanical jig, GOA/statewide)
top_permits$new[which(substr(top_permits$new,1,3)=="M26")] = "M26" # mechanical jig

# 6. Combine some of the longline categories based on vessel size.
top_permits$new[top_permits$new=="C06B"] = "C61B"
top_permits$new[top_permits$new=="B06B"] = "B61B"
top_permits$new[top_permits$new=="S05B"] = "S15B"

# 7. Combining salmon permits is a little more difficult, because the species composition
# varies widely based on geography. For the drift gillnet (S03) permits, we
# combined S03 permits from Cook Inlet, Bristol Bay, and Alaska Peninsula because sockeye
# represents > 92% of species revenue
top_permits$new[which(top_permits$new%in%c("S03H","S03M","S03T"))] = "S03" # drift gillnet
# For the purse seine permits (S01), again there's wide variety in species landings by
# area (85% pink in PWS to 85% sockeye on alaska peninsula). Of the 5 permits, only 2 had
# similar species compositions, so we grouped Kodiak and Cook Inlet
top_permits$new[which(top_permits$new%in%c("S01H","S01K"))] = "S01" # purse seine
# For the set gillnet, S04 permits, there's some variety but also clear sockeye specialists. We
# grouped permits that had > 80% sockeye
top_permits$new[which(top_permits$new%in%c("S04E","S04H","S04K","S04M","S04T"))] = "S04a" # purse seine
# We grouped permits from norton sound and kuskokwim because of similar species %
top_permits$new[which(top_permits$new%in%c("S04W","S04Z"))] = "S04b"
# There's a few remaining permits (S04D, S04X, S04P, S04Y) - but they're so different they can't be grouped

#8. combine the new permit groupings into new strategies
top_permits$orig = as.character(top_permits$orig)
top_strategies$new.strategy = NA
for(i in 1:nrow(top_strategies)) {
  top_strategies$new.strategy[i] = paste(top_permits$new[match(lapply(lapply(top_strategies$strategy, strsplit, " "), unlist)[[i]],
  top_permits$orig)], collapse=" ")
}
# join this into the data frame -- this is leaving ~ 70 strategies
dat = left_join(dat, top_strategies) %>% select(-c(n, earn))
dat$strategy=dat$new.strategy
dat = dat[is.na(dat$strategy)==FALSE,]

# 9. Remove data points where people only did a strategy in one year. Below
# we try to include people-strategy random effects
dat = group_by(dat, strategy, p_holder) %>%
  mutate(nsp = n()) %>% filter(nsp > 1) %>% select(-nsp)

# Derived variables
scale2 <- function(x) {
  x <- x - mean(x, na.rm = TRUE)
  x / (2 * sd(x, na.rm = TRUE))
}
dat$log_spec_div <- scale2(log(dat$specDiv))
dat$scaled_spec_div <- scale2(dat$specDiv)
dat$log_length <- scale2(log(dat$length + 1))
dat$log_weight <- scale2(log(dat$weight + 1))
dat$log_days <- scale2(log(dat$days + 1))
dat$log_npermit <- scale2(log(dat$npermit))
dat$log_days_permit <- scale2(log(dat$days_permit+1))

# delete strategy: region combinations with less than 50 observations
dat$strategy_region = paste(dat$strategy, dat$region)
dat = group_by(dat, strategy_region) %>% 
  mutate(n = n()) %>% 
  filter(n > 50) %>% select(-n)

# delete strategies in one region
dat = group_by(dat, strategy) %>% 
  mutate(n = length(unique(region))) %>% 
  filter(n > 1) %>% select(-n)

# drop regions with < 3 strategies
dat = group_by(dat, region) %>% 
  mutate(n = length(unique(strategy))) %>% 
  filter(n > 2) %>% select(-n)

dat$logdiff = log(dat$revenue/dat$revenue.prev)
dat$specdiff = log(dat$specDiv/dat$specdiv.prev)
dat$daysdiff = log((dat$days_permit+1)/(dat$days_permit.prev+1))
#mu.mod = glmmTMB(log(revenue) ~ specdiff * daysdiff + (1 + specdiff * daysdiff| strategy) + offset(log(revenue.prev)), data = dat)
m = ceiling(table(dat$strategy, dat$region)/1.0e10)
colnames(m) = c("AP", "Bering","BB","CookInlt","Kodiak","PWS","SEAK","Yakutat")
kable(m)
```

@@

```{r, fig.cap="Species diversity by year\\label{fig:diversitybyyear}", fig.pos="placeHere", echo=FALSE, eval=FALSE}
#p1 = ggplot(agg, aes(year, specDiv, colour = strategy_permit)) + geom_line() + ylab("Species diversity")
ggplot(agg.annual, aes(year, specDiv, colour = specDiv)) + geom_line() + facet_wrap(~ strategy_permit, scale="free") + ylab("Species diversity")
```

```{r, fig.cap="Mean revenue by year\\label{fig:medianrevenuebyyear}", fig.pos="placeHere", echo=FALSE, eval=FALSE}
#p1 = ggplot(agg, aes(year, log(meanrev), color = strategy_permit)) + ylab("Ln median revenue") + geom_line()
ggplot(agg.annual, aes(year, log(meanrev))) + facet_wrap(~ strategy_permit, scale="free") + ylab("Ln median revenue") + geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + geom_line()
```

@@

```{r, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE, eval=FALSE}
# We can also do the same calculations by calculating means for each year and then averaging across years, 
agg = group_by(dat, strategy_permit, year) %>% 
  summarize(specDiv = mean(specDiv), meanrev = mean(revenue),
    cv = sd(log(revenue/revenue.prev))) %>% 
  group_by(strategy_permit) %>%
  summarize(specDiv = mean(specDiv,na.rm=T), meanrev=mean(meanrev,na.rm=T),
    cv = mean(cv,na.rm=T))

p1 = ggplot(agg, aes(specDiv, log(meanrev), colour = cv, label=strategy_permit)) + xlab("Species diversity") +ylab("Ln (mean revenue)") + geom_text()
p2 = ggplot(agg, aes(log(meanrev), cv, colour = specDiv, label=strategy_permit)) + ylab("CV") + xlab("Ln (mean revenue)") + geom_text()
p3 = ggplot(agg, aes(specDiv, cv, colour = log(meanrev), label=strategy_permit)) + xlab("Species diversity") +ylab("CV") + geom_text()
gridExtra::grid.arrange(p1,p2,p3, ncol=2)
```

```{r, fig.cap="CV of revenue by year", fig.pos="placeHere", echo=FALSE, eval=FALSE}
#p1 = group_by(dat.diff, strategy_permit, year) %>% 
#  summarize(sd_logdiff = sd(log(revenue/revenue.prev))) %>% 
#  ggplot(aes(year, sd_logdiff, colour = strategy_permit)) + geom_line() + ylab("CV revenue")
p2 = group_by(dat.diff, strategy_permit, year) %>% 
  summarize(CV = sd(log(revenue/revenue.prev))) %>% 
  ggplot(aes(year, CV, colour = CV)) + geom_line() + 
  facet_wrap(~ strategy_permit, scale="free") + ylab("CV revenue")
p2
```

```{r, fig.cap="Mean revenue versus CV", fig.pos="placeHere", echo=FALSE, eval=FALSE}
group_by(dat.diff, strategy_permit, year) %>% 
  summarize(meanrev = mean(revenue), sd_logdiff = sd(log(revenue/revenue.prev))) %>% 
  ggplot(aes(log(meanrev), sd_logdiff, colour = sd_logdiff)) + 
  geom_point() + facet_wrap(~ strategy_permit, scale="free") + 
  ylab("CV revenue") + xlab("Ln mean revenue") + geom_smooth()
```

```{r, fig.cap="Species diversity versus ln (mean revenue)", fig.pos="placeHere", echo=FALSE, eval=FALSE}
group_by(dat.diff, strategy_permit, year) %>% 
  summarize(specDiv = mean(specDiv), meanrev = mean(revenue)) %>% 
  ggplot(aes(specDiv, log(meanrev), colour = meanrev)) + 
  geom_point() + facet_wrap(~ strategy_permit, scale="free") + 
  ylab("Ln (mean revenue)") + xlab("Mean species diversity") + geom_smooth()
```

```{r, fig.cap="Species diversity versus CV", fig.pos="placeHere", echo=FALSE, eval=FALSE}
group_by(dat.diff, strategy_permit, year) %>% 
  summarize(specDiv = mean(specDiv), sd_logdiff = sd(log(revenue/revenue.prev))) %>% 
  ggplot(aes(specDiv, sd_logdiff, colour = sd_logdiff)) + 
  geom_point() + facet_wrap(~ strategy_permit, scale="free") + 
  ylab("CV revenue") + xlab("Mean species diversity") + geom_smooth()
```

@@

###This plot shows the time series of becoming increasingly diversified from 1 species on mean revenue. For S01E for example, there's huge benefits of diversifying in 1991, and huge costs in 2010. 2010 was a huge pink salmon year for PWS (with high prices specialization was good) and 1991 was the start of pink salmon stock + price collapse. Perhaps slight declines in some of these in recent years?

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification by permit / strategy."}
ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0)) + ggtitle("Effect on Mean")
```

```{r, echo=FALSE}
df.cv = ranef(cv.mod)$cond$year_permit
colnames(df.cv) = c("c.cv","b.cv","a.cv")
df.cv$year = as.numeric(substr(rownames(df.cv), 1, 4))
df.cv$strategy = substr(rownames(df.cv), 6, 9)
# also calculate change of going from 1 to 2 spp
df.cv$pred1.cv = df.cv$a.cv*(1^2) + df.cv$b.cv*1 + df.cv$c.cv + fixef(cv.mod)$cond["specDiv.change"] + fixef(cv.mod)$cond["specDiv.change:specdiv.prev"]*1

#ggplot(df.cv, aes(year, pred1.cv, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0)) + ggtitle("Effect on CV")

@@

#df = left_join(df,df.cv)
#ggplot(df, aes(pred1, pred1.cv, label=year)) + facet_wrap(~strategy, scale="free")+ #geom_vline(aes(xintercept=0)) + geom_hline(aes(yintercept=0)) + xlab("Benefit on the mean") + #ylab("Benefit on the CV") + geom_text(size=2)
```

@@

```{r, echo = FALSE, fig.cap="Results for PWS Purse seine fishery", echo=FALSE, fig.pos="placeHere", cache=TRUE}

@@

```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", cache=TRUE, echo=FALSE, fig.pos="placeHere"}

@@

```{r, echo = FALSE, fig.cap="SEAK gillnet (clustering primarily in chum-sockeye-chinokook/coho dimensions)", echo=FALSE, fig.pos="placeHere", cache=TRUE, eval=FALSE}
fishery = "S03A"
df = dat[which(dat$strategy_permit==fishery),]
pamk_s03a <- pamk(df[,c("chum","pink","sock","salm")], 3)
df$group = pamk_s03a$pamobject$clustering

p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)

```{r, echo = FALSE, fig.cap="SEAK gillnet fishery (clustering primarily in sockeye-chinook/coho dimension)", echo=FALSE, fig.pos="placeHere", cache=TRUE, eval=FALSE}
fishery = "S03E"
df = dat[which(dat$strategy_permit==fishery),]
pamk_s03e <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s03e$pamobject$clustering

p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)

```{r, echo = FALSE, fig.cap="Kodiak purse seine (clustering primarily in sockeye-pink dimension)", echo=FALSE, fig.pos="placeHere", eval=FALSE, cache=TRUE}
fishery = "S01K"
df = dat[which(dat$strategy_permit==fishery),]
pamk_s01k <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s01k$pamobject$clustering

p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)

```{r, echo = FALSE, fig.cap="Kodiak gillnet (clustering primarily in sockeye-pink dimension)", echo=FALSE, fig.pos="placeHere", cache=TRUE, eval=FALSE}
fishery = "S04K"
df = dat[which(dat$strategy_permit==fishery),]
pamk_s04k <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s04k$pamobject$clustering

p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)

```{r, echo = FALSE, fig.cap="SEAK Purse seine (clustering primarily in sockeye-pink dimension)", echo=FALSE, fig.pos="placeHere", eval=FALSE, cache=TRUE}
fishery = "S01M"
df = dat[which(dat$strategy_permit==fishery),]
pamk_s01m <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s01m$pamobject$clustering

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)

@@

```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE}
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = mean(g_price)) %>% 
  group_by(spec, area) %>% 
  mutate(p = p/p[which(year==2002)]) %>% 
  group_by(year, area) %>%
  ggplot(aes(year, log(p), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
  ylab("Ln avg price relative to 2002")
```

```{r, fig.cap="Number of people landing fish", fig.pos="placeHere", echo=FALSE, eval=FALSE}
#Participation (people landing fish on one of these permits) generally followed trends in prices, with an uptick since early 2000s,  
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, p_fshy) %>%
  summarize(people = length(unique(p_holder))) %>%
  ggplot(aes(year, people)) + geom_line() + facet_wrap(~p_fshy, scale="free") + 
  ylab("People")
```

```{r, fig.cap="Trends in revenue and variability across people", fig.pos="placeHere", echo=FALSE}
#p1 = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
#  summarize(gearn = sum(g_earn)) %>%
#  group_by(year, p_fshy) %>%
#  summarize(cv = sd(gearn)/mean(gearn)) %>%
#  ggplot(aes(year, cv, color = p_fshy)) + geom_line() + 
#  ylab("CV revenue (across people")

p2 = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
  summarize(gearn = sum(g_earn)) %>%
  group_by(year, p_fshy) %>%
  summarize(med = median(gearn), low = quantile(gearn, 0.25),
    upp = quantile(gearn,0.75)) %>%
  ggplot(aes(year, log(med))) + geom_line() + facet_wrap(~p_fshy, scale="free") + 
  geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + ylab("Ln quartiles of revenue (across people)")
print(p2)
```

Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.  
```{r}
group_by(cfec, p_fshy, year, p_holder) %>% 
  summarize(meanW = mean(g_pounds/2000, na.rm=T)) %>% 
  group_by(p_fshy, year) %>% 
  summarize(meanW = mean(meanW, na.rm=T)) %>% 
  ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons per person") + xlab("Year")

```{r, fig.cap="CV of average species composition. This quantifies the synchrony across people -- so low CVs mean everyone's doing the same thing.", fig.pos="placeHere", echo=FALSE, eval=FALSE}
# Another way to think about diversification is to calculate the deviations in species composition from the mean. Bascially what's the variability of euclidian distances? Given trends in total landings and diversity, the 'portfolio' of species available to any of these permit holders has changed through time. Another way to think about the diversification is whether the variability of people around the mean has changed (basically whether there's been trends in the CV of species diveristy across people).

dat.salm = group_by(dat, year, p_holder) %>% 
  mutate(earn = (chnk+coho+chum+pink+sock),
    chnk=chnk/earn, coho=coho/earn,
    chum=chum/earn, sock=sock/earn,
    pink=pink/earn) %>% 
  as.data.frame 
dat.salm$chnk[is.na(dat.salm$chnk)] = 0
dat.salm$coho[is.na(dat.salm$coho)] = 0
dat.salm$chum[is.na(dat.salm$chum)] = 0
dat.salm$pink[is.na(dat.salm$pink)] = 0
dat.salm$sock[is.na(dat.salm$sock)] = 0

dat.salm = group_by(dat.salm, year, strategy_permit) %>% 
  mutate(mean.chnk = mean(chnk,na.rm=T), mean.chum = mean(chum,na.rm=T),
    mean.pink = mean(pink,na.rm=T), mean.coho = mean(coho,na.rm=T), 
    mean.sock = mean(sock,na.rm=T))

dat.salm$eucld = (dat.salm$mean.chnk-dat.salm$chnk)^2 + (dat.salm$mean.coho-dat.salm$coho)^2 + (dat.salm$mean.pink-dat.salm$pink)^2 + (dat.salm$mean.sock-dat.salm$sock)^2 + (dat.salm$mean.chum-dat.salm$chum)^2

p1 = group_by(dat.salm, year, strategy_permit) %>%
  summarize(sd = sd(eucld)) %>%
  ggplot(aes(year, sd, color = as.factor(strategy_permit))) + geom_line() + ylab("SD of euclidian dist.") + facet_wrap(~strategy_permit)

group_by(dat, year, strategy_permit) %>% 
  summarize(cv = sd(specDiv)/mean(specDiv)) %>% 
  ggplot(aes(year, cv, color = as.factor(strategy_permit))) + geom_line() + ylab("CV of species diversity (across people)") + facet_wrap(~ strategy_permit, scale="free") + geom_smooth()
```

>>>>

@@

@@

@@

###This plot shows the time series of becoming increasingly diversified from 1 species on mean revenue. For S01E for example, there's huge benefits of diversifying in 1991, and huge costs in 2010. 2010 was a huge pink salmon year for PWS (with high prices specialization was good) and 1991 was the start of pink salmon stock + price collapse. Across permits, perhaps slight declines in some of these in recent years?

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification by permit / strategy.", fig.height = 3}
ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + ggtitle("Effect on Mean")
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification split out by permit / strategy.", fig.height = 3}
ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0))
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of effects of diversification on mean and CV by permit."}
df.cv = ranef(cv.mod)$cond$year_permit
colnames(df.cv) = c("c.cv","b.cv","a.cv")
df.cv$year = as.numeric(substr(rownames(df.cv), 1, 4))
df.cv$strategy = substr(rownames(df.cv), 6, 9)
# also calculate change of going from 1 to 2 spp
df.cv$pred1.cv = df.cv$a.cv*(1^2) + df.cv$b.cv*1 + df.cv$c.cv + fixef(cv.mod)$cond["specDiv.change"] + fixef(cv.mod)$cond["specDiv.change:specdiv.prev"]*1

@@

@@

```{r, echo = FALSE, fig.cap="Results for PWS Purse seine fishery", fig.pos="placeHere", cache=TRUE}

@@

```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", cache=TRUE, fig.pos="placeHere"}

@@

@@

```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE, fig.height = 3.5}
g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = mean(g_price))
p1 = group_by(g, spec, area) %>% 
  mutate(p = p/p[which(year==2002)]) %>% 
  group_by(year, area) %>%
  ggplot(aes(year, log(p), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
  ylab("Ln avg price relative to 2002") + ggtitle("Synchrony in prices: relative change")
p1
```

```{r, fig.cap="Percent change (log-difference) in inflation adjusted prices / species by area", fig.pos="placeHere", echo=FALSE, fig.height = 3.5}
g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = mean(g_price))
# do log-diff of prices -- get price year before
g$p_t1 = g$p[match(paste(g$year, g$spec, g$area), 
  paste(g$year+1, g$spec, g$area))]
p2 = group_by(g, spec, area) %>% 
  ggplot(aes(year, log(p/p_t1), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
  ylab("Ln price (t/t-1)") + ggtitle("Synchrony in prices: percent change") 
p2
```

```{r, fig.cap="Trends in revenue and variability across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.5}
p2 = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
  summarize(gearn = sum(g_earn)) %>%
  group_by(year, p_fshy) %>%
  summarize(med = median(gearn), low = quantile(gearn, 0.25),
    upp = quantile(gearn,0.75)) %>%
  ggplot(aes(year, log(med))) + geom_line() + facet_wrap(~p_fshy, scale="free") + 
  geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + ylab("Ln quartiles of revenue (across people)") + ggtitle("Trends in revenue")
print(p2)
```

Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.  
```{r, echo=FALSE, fig.height = 3.5, fig.cap="Trends in average landed weight per person", fig.pos="placeHere"}
group_by(cfec, p_fshy, year, p_holder) %>% 
  summarize(meanW = mean(g_pounds/2000, na.rm=T)) %>% 
  group_by(p_fshy, year) %>% 
  summarize(meanW = mean(meanW, na.rm=T)) %>% 
  ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons per person") + xlab("Year") + ggtitle("Trends in landed weight")
```
